---
title: "Final Project Analysis"
author: "AHeredia"
date: "2022-11-29"
output: html_document
---
Installing packages and libraries
```{r}
pacman::p_load(haven, readr, dplyr, openxlsx, readxl, ggplot2, tidyverse, table1, tidyr, aod, DiagrammeR, finalfit, odds.n.ends,readr, broom, car)
```

Read data sets
```{r}
Demo <- read.csv("C:/Users/andre/Desktop/WashU/WashU 2nd Year/Fall 2022/Advanced Data Analysis/Final Project/RQ and Datasets identified/Demo.csv")

Ferritin <- read.csv("C:/Users/andre/Desktop/WashU/WashU 2nd Year/Fall 2022/Advanced Data Analysis/Final Project/RQ and Datasets identified/Ferritin.csv")

IndSupp <- read.csv("C:/Users/andre/Desktop/WashU/WashU 2nd Year/Fall 2022/Advanced Data Analysis/Final Project/RQ and Datasets identified/IndSupp.csv")
```

Merge data sets (DFI = Demographics Ferritin Individual Supplements)
```{r}
DemoFe <- merge(Demo,Ferritin, by="ID") ##merging demographic data with Ferritin levels data

DFI <- merge(DemoFe, IndSupp, by="ID") ##merging new data frame (Demo+Ferritin) with Individual Supplement data

remove(Demo,Ferritin, IndSupp, DemoFe) ##remove data sets that are not needed anymore
```

Keeping only variables to be used in analysis
```{r}
colnames(DFI)

DFI<-rename(DFI, c("FeLevels" = "LBDFERSI..ug.L.",
                 "Age" = "RIDAGEYR",
                 "Gender" = "RIAGENDR",
                 "VitC" = "DSQIVC",
                 "Iron" = "DSQIIRON"
                 ))

DataSet1 <- select(DFI, "ID","Age", "Gender", "FeLevels", "VitC", "Iron")#keeping only variables that will be used in analysis

remove(DFI) ##remove data set that is not needed anymore
```

Removing cases with NA's
```{r}
summary(DataSet1)
#FeLevels: 1011 NA'S
#VitC: 10073 NA's
#Iron: 12997 NA's
DataSet2<- drop_na(DataSet1, "FeLevels","Iron")##no NA's 

#only dropping NA's in Iron and FeLevels . FeLevels is the outcome variable and Iron the main exposure, which will be moderated by age, gender and vitC (main moderator of interest). NA's in VitC will be coded as 0 = people do not take vitc OR did not provide info
```


Logistic regression to see if taking iron supplements can predict if someone will be iron deficient
```{r}
summary(DataSet2)
#n=1573

#recoding outcome variables 0= >15ug/L = "Non-iron deficient" // 1= <15ug/L = "Iron-deficient"
DataSet2 <- DataSet2%>%
  mutate(FeLevels = case_when(FeLevels >= 15  ~ 0,
                          FeLevels < 14.9 ~ 1 ),
         FeLevels_Cat = factor(FeLevels, 0:1, c("Non-Iron-Deficient", "Iron-Deficient"))) 

summary(DataSet2)#2 NA's show up on FeLevels, so we proceed to remove them
DataSet3<-drop_na(DataSet2, "FeLevels")
summary(DataSet3)

#convert FeLevels to factor "FeLevels_Rec"
DataSet3$FeLevels_Rec <- factor(DataSet3$FeLevels)


#integers as numeric, and recoding gender
DataSet3$ID<-as.numeric(DataSet3$ID)
DataSet3$Gender<-as.numeric(DataSet3$Gender)
DataSet3$Age<-as.numeric(DataSet3$Age)
DataSet3$Gen_Rec<-as.numeric(DataSet3$Gender)

DataSet3 <- DataSet3 %>%
  mutate(Gen_Rec = case_when(Gen_Rec == 1 ~ 0,
                             Gen_Rec == 2 ~ 1),
         Gen_Cat = factor(Gen_Rec, 0:1, c("Male", "Female")))#Males are 0; Females are 1

#gender_rec as factor for analysis
DataSet3$Gen_Rec <- factor(DataSet3$Gen_Rec)

summary(DataSet3)
#recoding age variable to 5 different categories
DataSet3 <- DataSet3 %>%
  mutate(Age_Rec = case_when(Age > 0 & Age < 10 ~ 0,
                             Age >= 11 & Age < 20 ~ 1,
                             Age >= 21 & Age < 30 ~ 2,
                             Age >= 31 & Age < 40 ~ 3,
                             Age >= 41 & Age < 50 ~ 4,
                             Age >= 50 ~ 5),
         Age_Cat = factor(Age_Rec, 0:5, c("1-10", "11-20", "21-30", "31- 40", "41-50", ">50"))) #make it as a factor variable

#Age_rec as factor for analysis
DataSet3$Age_Rec <- factor(DataSet3$Age_Rec)

summary(DataSet3)
#recoding VitC variable to 2 categories
DataSet3$VitC[is.na(DataSet3$VitC)] = 0 #converting NA's to 0's

DataSet3 <- DataSet3 %>%
  mutate(VitC_Rec = case_when(VitC == 0 ~ 0,
                          VitC > 0 ~ 1),
        VitC_Cat = factor(VitC_Rec, 0:1, c("Does Not Take VitC", "Takes VitC"))) #make it as a factor variable >0=takes VitC; 1=Does NOT take VitC

#VitC_Rec as factor for analysis
DataSet3$VitC_Rec <- factor(DataSet3$VitC_Rec)
```

```{r}
# visualize FeLevels use by Iron using a box plots
summary(DataSet3)
# a 
YN<-DataSet3 %>% drop_na(FeLevels_Rec) %>%
  ggplot() +
  geom_boxplot(aes(x=Iron, y=FeLevels_Rec, fill=Iron)) +
  theme_bw() + xlab("Iron Supplement") + ylab("Iron-deficient (No/Yes)")
YN
# There should be a positive association between Iron and Ferritin Levels.
by(DataSet3$Iron,DataSet3$FeLevels_Rec, summary)
```

```{r}
# box tidwell technique, test the assumption of linearity
DataSet3$Iron_times_log_bmi<-DataSet3$Iron *log(DataSet3$Iron)
TestModel<-glm(FeLevels_Rec ~ Iron + Iron_times_log_bmi, data=DataSet3, family="binomial")
summary(TestModel)
# Since the p-value for Iron_times_log_bmi <.05. we conclude the assumption is violated.If we take 0.05 as our cutoff threshold
```
```{r}
####check classification
table(DataSet3$FeLevels_Rec, DataSet3$FeLevels, useNA="always")
```
Running an univariate logistic regression of the association between FeLevels category and `Iron` and output a summary using the summary function.**
```{r, echo=FALSE, eval=FALSE}
LR1<-glm(FeLevels ~ Iron, data=DataSet3, family="binomial") 
summary(LR1)
odds.n.ends(LR1)

###Interpretation: for every unit increase in Iron, the odds of being iron deficient increase by 1.2% (95%CI 1.01-1.02)
#as iron levels go up the odds of being iron deficient go up. 
```
Logistic regression 2 just with iron and gender recoded. Reference group is 0 (males and non-iron deficient) reference group is lowest coding category
```{r}
LR2 <- glm(FeLevels ~ Iron + Gen_Rec, data = DataSet3, family = "binomial")
summary(LR2)
confint(LR2)
exp(cbind(OR = coef(LR2), confint(LR2)))

odds.n.ends(LR2)
#controlling for iron levels, women relative to men have 8.17 times the odds of being iron deficient
```
Logistic regression 3 with iron, gender recoded and age recoded
```{r}
LR3 <- glm(FeLevels ~ Iron + Gen_Rec + Age_Rec, data = DataSet3, family = "binomial")
summary(LR3)
confint(LR3)
exp(cbind(OR = coef(LR3), confint(LR3)))
odds.n.ends(LR3)
##compared to those who are 1-10 years old the odds of being iron deficient are 5.43 times higher than those who are 11-20

```
Logistic regression 4  with iron,gender recoded, age recoded, and Vitamin C recoded
```{r}
LR4 <- glm(FeLevels ~ Iron + Gen_Rec + Age_Rec + VitC_Rec, data = DataSet3, family = "binomial")
summary(LR4)
confint(LR4)
exp(cbind(OR = coef(LR4), confint(LR4)))
odds.n.ends(LR4)
###After adjusting for all the variables the total effect of iron on felevels is no longer significant. 
```
Testing interaction between gender and age: does the association we see on iron and felevels vary on gender? is taking iron different for men? or women? no difference?
```{r}
LR5 <- glm(FeLevels ~ Iron*Gen_Rec + Age_Rec + VitC_Rec, data = DataSet3, family = "binomial")
summary(LR5)
odds.n.ends(LR5)
#the effect on iron does not depend on gender
```

Testing interaction between gender, age, and vitamin C intake: does the association we see on iron and felevels vary on gender? on gender and age? on gender, age, and vitamin C intake? no difference?
```{r}
LR6 <- glm(FeLevels ~ Iron + Gen_Rec*Age_Rec*VitC_Rec, data = DataSet3, family = "binomial")
summary(LR6)
odds.n.ends(LR6)
```

```{r}
# Compare the coefficients between models with and without these observations
compareCoefs(LR1, LR2, LR3, LR4, LR5, LR6)
```
Probability plots
```{r}
odds.n.ends(LR1, predProbPlot=TRUE, rocPlot=TRUE)
odds.n.ends(LR2, predProbPlot=TRUE, rocPlot=TRUE)
odds.n.ends(LR3, predProbPlot=TRUE, rocPlot=TRUE)
odds.n.ends(LR4, predProbPlot=TRUE, rocPlot=TRUE)
odds.n.ends(LR5, predProbPlot=TRUE, rocPlot=TRUE)
odds.n.ends(LR6, predProbPlot=TRUE, rocPlot=TRUE)
```
Exporting dataset 
```{r}
write.csv(DataSet3,"C:/Users/andre/Desktop/WashU/WashU 2nd Year/Fall 2022/Advanced Data Analysis/Final Project/Abstract/DataSet.csv", row.names = FALSE)
```

Data visualization
```{r}
#Data visualization
label(DataSet3$Gen_Cat)<-"Gender"
label(DataSet3$Age_Cat)<-"Age (Years)"
label(DataSet3$VitC_Cat)<-"Vitamin C"

table1(~ Iron + Gen_Cat + Age_Cat + VitC_Cat | FeLevels_Cat, overall="Total",rowlabelhead="Variable",footnote='Reference categories = Male, 1-10, and Does Not Take VitC', data=DataSet3)



#PLOTTING LOGISTIC REGRESSION CURVE
ggplot(DataSet3, aes(x=Iron, y=FeLevels)) +
  geom_point(alpha=.5) +
  stat_smooth(method="glm", se=T, method.args = list(family=binomial),
              col="red", lty=2)

table(DataSet3$FeLevels_Rec, useNA = 'ifany')
```
```{r}
library(DiagrammeR)
#Figure 1
grViz("digraph flowchart {

      # node definitions with substituted label text
      node [fontname = Helvetica, shape = rectangle, fontsize=8] 
      # notes indicate how many boxes you will have in your diagram. Since I have two sets of exclusions, I will have three nodes.
      tab1 [label = '@@1'] # starting
      tab2 [label = '@@2'] # exclusion 1
     
      
      # edge definitions with the node IDs are used to indicate how the rectangle boxes flow from each other. This is a simple flow chart but I could create branching instead as below
      tab1 -> tab2 
}
      #This set of code provides the text in each rectangle box.
      [1]: 'Secondary data from NHANES for people whose Ferritin levels were measured, had taken Iron and Vitamin C as oral supplements n=14,696'
      [2]: 'Excluding 13,123 individuals who had missing data on the Ferritin levels, and Iron columns n=1,571'
      ")
```